---
title: "ETC5550 Assignment 2"
author: "Arindom Baruah (32779267)"
date: "2024-03-22"
quarto-required: ">=1.3.0"
format:
    html:
        output-file: assign02-submission.html
        css: "assignment.css"
execute: 
  echo: false
  message: false
  warning: false
number-sections: true
---

# Libraries used

```{r}
#| echo: TRUE
library(fpp3)
library(tidyverse)
library(kableExtra)
```

# Glimpse of the data

```{r}
#| echo: false
get_my_data <- function(student_id) {
  set.seed(student_id)
  all_data <- fill_gaps(readr::read_rds("https://bit.ly/monashretaildata"))
  while(TRUE) {
    retail <- filter(all_data, `Series ID` == sample(`Series ID`, 1)) 
    if(!any(is.na(retail$Turnover))) return(retail)
  }
}
# Replace the argument with your student ID
retail <- get_my_data(32779267)
```


Once we have obtained our data, let us take a quick glimpse of the data as illustrated by @tbl-data. 

```{r}
#| label: tbl-data
#| tbl-cap: "Sample of the current retail data"

retail %>% head() %>% kbl()
```

# Exercises

## Plotting of timeseries

Let us observe the time series data for the turnover generated.

```{r}
#| label: fig-turnover
#| fig-cap: "Retailing turnover in Australia over time"
retail %>% autoplot(Turnover) + labs(x = "Date",y = "Turnover (mil $)",title="Retailing turnover in Australia over time",caption = "Source: Australian Bureau of Statistics") + theme_minimal()
```
:::{.callout-note}

Based on @fig-turnover we can observe the following details in the timeseries plot:

1. There appears to be a seasonality in the plot with a spike being observed on every January of a year. __The seasonality additionally appears to be consistent throughout the plot.__

2. The amplitude of the spikes observed in the data appear to be increasing with time. __This indicates that the variation in the data is multiplicative in nature.__

3. There appears to be an __outlier in the data during the year of 2020.__ While on each January, we expect a rise in turnover, however, in the year of January 2020, the turnover was considerably lower than all the other years. This is expected to be caused due to __shutting down of businesses and their activities in the initial part of the pandemic.__

:::